---
title: "Initial Korean covid 19"
author: "Caitlyn Hense"
date: "4/15/2025"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?






## Graphical results - use ggplot commands to do this  
```{r}
dd2 <- kRdata %>% group_by(result,sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=sex,y=prcnt*100,fill=result)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

There are more females who got COVID and lived than males. However, out of the ones who got COVID, and passed, the percentage of men who died is larger than the percent of women who died. Your chances of dying if you caught COVID as a male were higher than a female. 
  





  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

Out of the alive group. 12% more women survived over men. Out of the dead group, about 36% of men over women died. 98.7% of females lived, while 1.4% of then died. 97% of men lived while 3% of the men died. percentage wise, COVID killed about twice the amount of men than women. 


## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

In out Chi-Square result, we recieved a p-value of 0.0012. Since this value is <0.05 (our point of significant), we reject or null (that dying is not independent of sex). In our Fisher's test, we received a p-value of 0.0009 which is also <0.05, we again will be rejecting the null hypothesis.

## Conclusions

The question was if dying from COVID  is independent of sex. We used a sample from South Korea and then used graphical, numerical, and inferential results to research this question. From these, we discovered that 98.7% of females lived and 1.4% died compared to 97% of men living and 3% of them dying. percentage wise, COVID killed over twice the amount of men died than women. Eventually, we decided to reject or null hypothesis (dying isn't independent of sex) because our p-values show less than the value of significance. 


