---
title: "Initial Korean covid 19"
author: "Caitlyn Hense"
date: "4/15/2025"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?






## Graphical results - use ggplot commands to do this  
```{r}
dd2 <- kRdata %>% group_by(result,sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=sex,y=prcnt*100,fill=result)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

There are more females who got COVID and lived than males. However, out of the ones who got COVID, and passed, the percentage of men who died is larger than the percent of women who died. Your chances of dying if you caught COVID as a male were higher than a female. 
  





  
## Numerical results


## Inferential results 




## Conclusions

